\input{c://Dropbox/mimoto/def_mimoto0.7}
%------------------------------------------------------------------- 

\begin{center} {\large \bf 489/689  Assignment 3}\\ \end{center}


{\large Due: \textcolor{red}{ Tue Oct 8th. }}

%{\large Name: $\underline{ \hs80mm }$ }

\textcolor{red}{ 
{\large This HW has been revised. Now have two questions. }
}

\vs5mm 

\noi Notation of this assignment follows that of Cryer. 
 i.e. ARMA($p,q$) parameters are defiend 
\ccc
\Phi(B) Y_t = \Th(B) e_t,
\ecc
where 
$\Th(x) = 1-\th_1x - \th_2 x^2 - \cdots - \th_q x^q.$
There fore when you use R function, you muse use 
[$\th_1$ in R] $ = - \th_1$.  

\vs5mm 
\noi Submit printout of the R code as well as your answer.  
Class web page has sample code for Monte Carlo Simulation of ARMA. 

\subsection*{Questions:}

%------------------------------------------------------------------- 
\begin{enumerate}


%----- 
\item 
Consider an ARMA(1,1) model with $\phi = 0.5$ and \textcolor{red}{ $\th = -0.45$ }, with 
mean of 3 and iid N(0,1) errors.  

\begin{enumerate} 
\item For n = 100, evaluate the variances and correlation of the maximum likelihood
estimators of $\ph$ and $\th$ using Equations (7.4.13) on page 161. 

\item Repeat part (a) but now with n = 300. 


\item By Monte Carlo Simulaiton, verify the convergence of variance and correlation of the 
MLE of $\ph$ and $\th$ to asymptotic values obtained in part (a) and (b) above. 

\vs5mm 
Monte Carlo Simulation can be performes as following:
\begin{enumerate} 
\item Simulate ARMA process with given specification. 
\item Obtain MLE for simulated series.  Store your $\h \ph$ and $\h \th$ 
in array. 
\item Repeat (i) and (ii) many times (say 1000 times).
\item Look at the distribution of your 1000 realization of $\h \ph$ and 
$\h \th$.  Compute sample variance and correlation from the 1000 pairs. 
\end{enumerate}

\item What does above Monte Carlo Simulaiton, indicate about the bias in 
$\h \ph$ and $\h \th$? 

\item What does your simulation indicate about using asymptotic property of 
MLE?

\end{enumerate}


\clearpage 
%----- 
\item 
Consider an ARMA(1,1) model with $\phi = 0.5$ and \textcolor{red}{ $\th = 0.45$ }, with 
mean of 3 and iid N(0,1) errors.  

\begin{enumerate} 
\item For n = 1000, evaluate the variances and correlation of the maximum likelihood
estimators of $\ph$ and $\th$ using Equations (7.4.13) on page 161. 

\item Repeat part (a) but now with n = 3000. 


\item By Monte Carlo Simulaiton, verify the convergence of variance and correlation of the 
MLE of $\ph$ and $\th$ to asymptotic values obtained in part (a) and (b) above. 


\item What does above Monte Carlo Simulaiton, indicate about the bias in 
$\h \ph$ and $\h \th$? 

\item What does your simulation indicate about using asymptotic property of 
MLE?

\end{enumerate}


\item Plot histgram of $\h \ph$ and $\h \th$ for 
n=100, 300 from Problem 1, and for n=1000, n=3000 from Problem 2.  
Use layout() to plot 4 histogram in one page. 


\begin{verbatim} 
---------------------
X <- rnorm(100)

layout(matrix(1:4, 2,2))
hist(X)
hist(X)
hist(X)
hist(X)
layout(matrix(1,1,1))
\end{verbatim}



\end{enumerate}


\end{document}
%------------------------------------------------------------------- 
%------------------------------------------------------------------- 
